### 2015-06-05 text/ms.Rmd
Shorten the abstract
Still too long, should be < 200 words

What can we do when multiple ecological models give different impressions of population status? Any single model is unlikely to perform best across all species, exploitation dynamics, and statuses. Ensemble methods from the fields climate science and machine learning provide a powerful tool to optimally leverage the information contained in multiple ecological models. Here we evaluate the potential for ensemble methods to improve estimates of fish population status. We fit four data-limited models of population and exploitation dynamics and combine their estimates of the mean and slope of biomass at maximum sustainable yield (\bbmsy\\) with four ensemble models: a simple average, a linear model, and two machine learning methods---random forests and boosted regression trees. We build our ensemble models on a fully factorial simulated dataset of 5760 fish stocks and test our models on a cross-validated version of this dataset and a global database of `r ram_stocks_n` assessed fish stocks. We find that ensemble models substantially improve estimates of population status and trajectory. Machine learning ensemble models performed the best: `r mean_sim$mach_vs_ind_mare_fold`% improved accuracy (reduction in median absolute proportional error), rank-order correlation improved from `r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median proportional error) declined from `r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. We found similar improvements in performance when predicting the slope of status. When the same simulation-trained ensemble models were applied to a separate dataset of global fish stocks and compared to mean status from data-rich stock assessments, the best ensembles still improved all dimensions of performance (accuracy, correlation, and bias) over the next best individual model. Ensemble models have great potential to improve estimates of population status and trajectory across many taxa beyond fish. However, ensemble methods must be carefully tested, formed from a diverse set of accurate models, and built on a dataset sufficiently representative of the populations they are applied to.

>>>>

What can we do when multiple ecological models give different impressions of population status? Any single model is unlikely to perform best in all scenarios. However, ensemble methods from the fields climate science and machine learning provide a powerful tool to leverage information across multiple ecological models. Here we evaluate the potential for ensemble methods to improve estimates of fish population status. We fit four data-limited models of population and exploitation dynamics and combine their estimates of the mean and slope of biomass at maximum sustainable yield (\bbmsy\\) with four ensemble models: a simple average, a linear model, and two machine-learning methods---random forests and boosted regression trees. We build our ensembles on a simulated dataset of 5760 fish stocks and test our models with cross-validation and a global database of `r ram_stocks_n` stock assessments. We find that ensemble models substantially improve estimates of population status and trajectory. Machine-learning ensembles performed the best: accuracy improved `r mean_sim$mach_vs_ind_mare_fold`%, correlation improved from `r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median proportional error) declined from `r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. We found similar improvements when predicting the slope of status. When the simulation-trained ensembles were applied to global fish stocks, the best ensembles still improved all dimensions of performance over the next best individual model. Ensemble models have potential to improve estimates of population status and trajectory across many taxa beyond fish. However, ensemble methods must be carefully tested, formed from a diverse set of accurate models, and built on a dataset sufficiently representative of the populations they are applied to.
